\scuba: iterative map-making with the Sub-Millimetre User Reduction Facility

### 2012-09-06 smurf_paper.tex
comments from Gary

\begin{abstract}
  The Submillimetre Common User Bolometer Array 2 (\scuba) is an
  instrument operating on the 15-m James Clerk Maxwell Telescope,
  consisting of about 5000 bolometers in each of two simultaneous
  imaging bands centered over 450 and 850\,\micron. The camera is
  scanned across the sky and records data at a rate of 200\,Hz. The
  data rate is much greater than that of previously existing
  submillimetre cameras, and represents a significant data anlaysis
  challenge. We describe the production of \scuba\ maps using the
  Sub-Millimetre User Reduction Facility (SMURF) in which we have
  adopted a fast, iterative approach to map-making that enables data
  reduction on single, modern, high-end desktop computers with
  execution times that are shorter than the observing times.  SMURF is
  used in an automated setting both at the telescope for real-time
  feedback to observers, as well as for the production of generic
  science products at the Canadian Astronomical Data Centre. A suite
  of tools are also provided for interactive data analysis.  Three
  detailed case studies are used to: (i) explore convergence
  properties of the map-maker using simple prior constraints (Uranus
  -- a point source); (ii) achieve the white-noise limit of the
  instrument for faint point-source studies (blank-field survey of the
  Lockman Hole); and (iii) demonstrate that our strategy is capable of
  recovering angular scales up to the size of the array footprint
  (approximately 5\,arcmin) for bright extended sources (star-forming
  region M17).
\end{abstract}

@@

%\textbf{Guideline for figures:}

%\begin{itemize}
%\item minimize excess whitespace around the figure to optimize usage of space
%\item ensure a decent line thickness
%\item use a Times font for labels with a thickness that resembles the
%  text in the body. The font size should be similar to the main body
%  text or larger
%\item avoid using colour unless it is necessary for emphasis
%\end{itemize}

%\textbf{start intro here...}

The Submillimetre Common User Bolometer Array 2
\citep[\scuba,][]{2006SPIE.6275E..45H,holland2012} is a new instrument
for the 15-m James Clerk Maxwell Telescope (JCMT) on Mauna Kea,
Hawai'i. The camera simultaneously images the sky in two broad bands
centered over 450 and 850\,\micron, with approximately 7.5 and
14.5\,arcsec full-width at half-maximum (FWHM) point spread functions
(PSFs). The focal planes at each wavelength are populated with 4
rectangular subarrays, consisting of $40 \times 32$ bolometers each,
and together subtend a nearly 7\,arcmin $\times$ 7\,arcmin footprint
on the sky (excluding gaps, the continuous solid angle is about
43\,arcmin$^{2}$ per focal plane).  The instrument is also equipped
with a Fourier transform spectrometer \citep[FTS-2,][]{gom2010}, and a
polarimeter \cite[POL-2,][]{2011ASPC..449...68B}. This paper describes
the properties of \scuba\ data that are relevant to producing maps of
the imaging data, and the Submillimetre User Reduction Facility,
SMURF, a software package for perfoming the reduction written using
the Starlink Software Environment \citep{2009ASPC..411..418J}. The
analysis of FTS-2 and POL-2 data will be described at a later date
once these additional instruments have been commissioned and
characterized. The details of the instrument design, performance, and
its calibration are given in two companion papers: \citet{holland2012}
and \citet{dempsey2012}.

@@

There are numerous ways to attack this map-making problem, both in
terms of the data-collection method, and processing. The two most
important principles to follow in terms of scan strategy are: (i) to
modulate the astronomical signals of interest in such a way that they
appear in the lowest-noise regions of the bolometer noise power
spectrum, i.e., above the $1/f$ knee; and (ii) to provide good
``cross-linking'', in which each portion of the map is sampled on a
range of temporal scales, again, to help distinguish time-varying
noise features from fixed astronomical sources. In the case of \scuba,
(i) is achieved through fast-scanning of the entire telescope (up to
600\,arcsec\,sec$^{-1}$ slew speeds), such that significant drift in
the bolometers due to low-frequency noise occurs more slowly than the
crossing times for the astronomical scales of interest; and (ii) by
offering scan patterns that cross the sky at a wide range of position
angles, ultimately observing each portion of the map at a number of
different times. Such methods are now used by virtually all existing
ground-based submillimetre cameras
\citep[e.g.,][]{glenn1998,weferling2002,wilson2008,kovacs2008b}., in
favour of ``chopping'' methods (where the secondary is quickly moved
to modulate the signal) that were more appropriate for older
instruments that had poorer low-frequency noise performance, and were
only sensivitve to modest angular scales

There are three general styles of map-making that are relevant to
reducing bolomter data in the literature. The simplest ``direct
methods'' involve some basic processing of the data to remove as much
noise contamination as possible, (e.g., using baseline removal and
other simple filters), and then re-gridding these cleaned data into a
map. Such was the basic recipe for the reduction of chopped data from
SCUBA-2's predecessor SCUBA
\citep{1998ASPC..145..216J,2000ASPC..216..559J}, and MAMBO
\citep[e.g.,][]{omont2001}, another camera from the same
generation. Generally speaking, such methods are fast, although
dependending on the science goals and noise properties of the data,
they may not achieve the best noise performance on the angular scales
of interest. In more recent years a method that has been popular for
reducing data in fields of faint point sources with BOLOCAM
\citep[e.g.,][]{laurent2005}, and its younger sibling AzTEC
\citep[e.g.,][]{scott2008}, is ``principal component analysis'' (PCA)
cleaning. A statistical black-box removes the most correlated
components of the bolometer signals, enabling the detection of
point-sources very close to the theoretical white-noise limits of the
detectors with reasonable computation times when small numbers of
bolometers are involved (hundreds rather than thousands of
detectors). However, PCA cleaning is not a good generic solution for
producing maps of extended structures, since such sources produce
correlated signals amongst many detectors, and are removed by this
procedure.

The best existing map-making strategies for recovering information on
all angular scales are \emph{maximum likelihood} techniques, in which
the time-series data are expressed as a sampling of the ``true'' map
of the sky plus noise, and then the equation is inverted to estimate
the map as some weighted linear combination of the data that minimizes
the variance.  The first good description of this method appears in
\citet{janssen1992} for the production of maps from the COsmic
Background Explorer (COBE -- the description is relevant despite the
fact that it used a differential radiometer instead of
bolometers). Other descriptions in the CMB literature include
\citet{tegmark1997} and \citet{stompor2002}, and an application to
data from SCUBA is described in \citet{borys2004}.  The downside to
such methods is that they can be both computationally expensive and
have large memory requirements. While for some experimental designs
fast iterative methods for the inversion do exist with minimal memory
usage \citep[e.g.,][]{wright1996}, for the more general map-making
problem involving many detectors, things are significantly complicated
both by the need to measure the cross power-spectra for all unique
pairs of bolometers, and then to perform the inversion itself. The
most promising maximum-likelihood method that may one day be applied
to \scuba\ data is ``SANEPIC'', which successfully reduced maps from
the Balloon-borne Large-Aperture Submillimeter Telescope, involving
hundreds of bolometers, and correctly incorporating inter-bolometer
noise correlations \citep[][]{patanchon2008}.

@@

The third approach adopted here for \scuba\ is a compromise between
the previous two methods. Under the assumption that a significant
portion of the (predominantly low-frequency) non-white noise sources
can be modeled, an iterative solution is obtained for both the
astronomical image and the parameters of the noise model. Since the
remaining (non-modelled) noise sources are assumed to be white, a
single scalar noise may be calculated for all of the data points from
a given bolometer (since the noise at any instant in time is
uncorrelated with others, nor with the data from other bolometers),
greatly simplifying the measurement of noise properties and the
inversion step that is so complicated in the maximum-likelihood
methods. In principle, it should also be possible to recover a wider
dynamic range in angular scales than the simplest ``direct'' methods,
with an increase in calculation time that is merely linear in the
number of iterations needed, while using a comparable amount of
memory. A fast iterative method for handling chopped data from CMB
experiments was described in \citet{wright1996}, and later applied to
chopped SCUBA data by \citet{johnstone2000}, although in each case the
chopping renders the data essentially white without the need to model
additional noise components. More similar to the algorithm described
here was the iterative pipeline recipe for fitting and removing
baseline drifts in SCUBA scan-map data as the astronomical image
estimate improved \citep{1999ASPC..172..171J}. The two closest modern
relatives of SMURF are the Comprehensive Reduction Utility for SHARC-2
\citep[CRUSH,][]{kovacs2008}, and the pipeline developed for the
Bolocam Galactic Plane Survey \citep{aguirre2010}.

A reasonable model for correlated noise in \scuba\ is a single
``common-mode'' signal seen by all of the bolometers. Iterative
estimation and removal of this signal significantly lowers the $1/f$
knee, without compromosing structures on angular scales smaller than
the array footprint. Residual independent drifts at lower frequencies
are removed with an iterative high-pass filter. This strategy enables
SMURF to reduce data faster than they are taken, on single high-end
desktop computers, and it has successfully been used as part of a
real-time pipeline offering feedback to observers at the
telescope. The pipeline is also used to generate near-science grade
products using the JCMT Science Archive \citep{2011ASPC..442..203E}
hosted by the Canadian Astronomical Data Centre (CADC). Details of the
pipeline design are given in \citet{gibb2005} and
\citet{2008AN....329..295C}.

This paper is organized in the following way. We first describe the
properties of \scuba\ data in Section~\ref{sec:data} to motivate the
data model that we have adopted. Next, the details of the SMURF
algorithm are given in Section~\ref{sec:algorithm}. The paper is
concluded in Section~\ref{sec:examples} with three detailed test cases
that span the majority of observation types likely to be undertaken
with \scuba, with an emphasis on the mitigation of divergence
problems, and characterizing the output maps: (i) Uranus, a bright,
compact source (Section~\ref{sec:point}); (ii) the Lockman Hole, a
blind survey of faint point-like source (Section~\ref{sec:cosmo}); and
(iii) the star-forming region M17, including bright, extended emission
(Section~\ref{sec:extended}). All of the data analyzed in this paper
are publicly available through the CADC \scuba\ raw-data queries page
for the dates and observation numbers given in the
text\footnote{\texttt{http://www3.cadc-ccda.hia-iha.nrc-cnrc.gc.ca/
jcmt/search/scuba2}}. All of the analysis was performed using
\texttt{Starlink} version ... .

@@

In this section we summarize how \scuba\ works,
(Section~\ref{sec:bolos}), give exampels of typical bolometers signals
(Section~\ref{sec:bolosignal}), and the impact of magnetic field
pickup (Section~\ref{sec:magpickup}), and finally illustrate the use
of principal component analysis to explore the correlated noise
properties of \scuba\ data (Section~\ref{sec:pca}).

@@

Incoming light passes through a beam-splitter, and then bandpass
filters, providing simultaneous illumination and wavelength definition
of both the 450 and 850\,\micron\ focal planes. Each focal plan is
populated by four \emph{subarrays} (labeled s4a--s4d at 450\,\micron,
and s8a--s8d at 850\,\micron), each consisting of 32 columns and 40
rows of bolometers. The bolometers are thermal absorbers coupled to
superconducting transition-edge sensors (TESs) for thermometry;
temperature variations in the TESs produce changing currents, and
therefore varying magnetic fields.  The changing magnetic fields are
detected and amplified using chains of superconducting quantum
interference devices (SQUIDs), and the larger output currents are
digitized. Each detector has its own SQUID for the first-stage of the
amplification, but the remaining stages occur within a common chain of
SQUIDs for each of the 32 columns. All 40 rows are read-out in
sequence at a row-visit frequency of about 12\,kHz. Such a high sample
rate is unnecessary to produce maps, so the data are re-sampled to
approximately 200\,Hz before writing to disk.  It should also be noted
that there is an additional 41st row of SQUID readouts that are not
connected to TESs. These \emph{dark SQUIDs} track non-thermal noise
sources that are common to each column's amplifier chain. The
relationship between the output digitized current, $I$, and the input
power, $P$, or $dI/dP$ is established using \emph{flatfield}
observations immediately prior to science observations, in which the
output signal is measured throughout a ramp of the pixel heaters
(which provide a known input power); see Section~2.1 in
\citet{dempsey2012} and Section~5.3 in \citet{holland2012}. Finally,
the conversion to astronomical flux units from power involves a
correction for atmospheric extinction \citep[primarily using the JCMT
Water Vapour Monitor to track line-of-sight opacity variations, see
Section~3 in][]{dempsey2012}, and the application of a flux conversion
factor (FCF) which is established from regular measurements of
calibrators such as Uranus \citep[Section~5 in][]{dempsey2012}.

@@

We note that the noise performace of \scuba\ has evolved over time. An
initial ``\scuba\ shared-risk observing'' period (S2SRO) took place
during February and March 2010, during which each of the 450 and
850\,\micron\ focal planes were populated with single subarrays, s4a
and s8d respectively. In addition to having significantly fewer
available bolometers than the current fully-commissioned instrument
(and therefore a mapping speed reduced by a factor of $\sim$4),
instabilities in the fridge lead to a large periodic signal with a
period of about 25\,s that was correlated amongst the detectors. After
upgrading and commissioning the instrument, a servo using newly added
NTD thermometers in the focal planes has effectively mitigated this
problem \citep[Section~2.5 in][]{holland2012}. In addition, there were
improvements to the magnetic shielding (reducing magnetic field pickup
as described in Section~\ref{sec:magpickup}), as well as more
effective removal of aliased noise sources (which has reduced both the
presence of line features and the mean white noise level).

In order to reduce the impact of noise on the final maps, it is
essential that an appropriate scan strategy be used. First, it is
clearly advantageous that the astronomical signals appear in the
lowest-noise region of the power spectrum, namely frequencies above
$f_\mathrm{knee}$. This is achieved by scanning the telescope at large
slew rates, which are practically limited to about
600\,arcsec\,s$^{-2}$ for the JCMT in order to have reasonable
turn-around overheads (the amount of time spent changing direction and
speed, over the largest elevation range possible). At this maximum
speed there is a sample every 3.0\,arcsec, or approximately one third
of the 450\,\micron\ diffraction-limited full-width at half-maximum
(FWHM) -- a typical rule-of-thumb for adequately sampling a Gaussian
point spread function (PSF). The scan \emph{patterns} are designed,
primarily, to mitigate the effects of low-frequency $1/f$ drift. The
simplest scan one might use is a raster ``boustrophedon'' pattern,
i.e., a series of parallel passes across the field. However, such a
scan yields extremely poor performace for structures that are
transverse to the scan direction, since a long time will have elapsed
between each parallel step. In addition, other periodic noise features
may conspire with such a regularly-spaced pattern to produce periodic
structures in the map, or \emph{scan-synchronous} noise.  It has
therefore been recognized that scan strategies that visit each region
of the map on as many different time scales, and scan angles as
possible, offer the best protection against large-scale noise
\citep[see discussion in][and references therein]{kovacs2008b}; such
strategies are said to offer good \emph{cross-linking}. As a tradeoff
between truly random scan patterns and feasibility of implementation
taking into consideration both mechanical and scheduling constraints,
\scuba\ presently has two primary scan modes: PONG, which covers
rectangular areas larger than the array footprint in a closed pattern
that ``bounces'' off the edges at 45\,degree angles (and also usually
combined with a rotation through a number of fixed position angles to
create a ``rotating PONG'' with even better cross-linking) which was
based on a similar scan mode developed at the Caltech Submillimeter
Observatory; and a ``constant velocity daisy'' (CV Daisy) for small
areas (of order the array footprint), in which the telescope moves in
a circle, whose centre also slowly traces out a circle. For a more
complete description of the \scuba\ observing modes, see Section~5 in
\citet{holland2012}.

\begin{figure*}
\includegraphics[width=0.49\linewidth]{bolos_point_mix_s2sro.pdf}
\includegraphics[width=0.49\linewidth]{bolos_point_mix.pdf}
\caption{A comparison between single bolometer time-series in each of
  the 450 and 850\,\micron\ bands with the mixing chamber temperature
  and azimuth/elevation pointing offsets, before and after upgrading
  the instrument. The grey signals overplotted in the top two panels
  show the residual time-series after removing the common-mode
  signals.  Left: data taken during the S2SRO period, observation 29
  on 20100313. There is a strong correlation between the bolometers
  and the roughly $\sim$25\,s oscillation in the fridge, but only a
  minor correlation with the telescope motion. The nearly flat
  common-mode subtracted signals show: (i) that most of the
  low-frequency signal is common to all of the bolometers; and (ii)
  the non-correlated, and predominantly white noise at 450\,\micron\
  is significantly larger than at 850\,\micron. Right: data taken with
  the fully-commissioned instrument, observation 38 on
  20111112. Unlike the S2SRO data, there is no strong signal produced
  by variations in the fridge temperature. The 450\,\micron\ data have
  minimal correlated low-frequency noise (as evidenced by the lack of
  a difference between the raw and common-mode subtracted signals),
  although the 850\,\micron\ data show a common signal correlated with
  the telescope motion (note that this scan has a significantly larger
  amplitude than the S2SRO data set), mostly likely caused by magnetic
  field pickup (see Section~\ref{sec:magpickup} and
  Fig.~\ref{fig:magpickup}).}
\label{fig:bolos_mix}
\end{figure*}

@@

Data from the fully commissioned instrument, observation 38 on
20111112, are shown in the right panel of
Fig.~\ref{fig:bolos_mix}. Having solved the fridge oscillation
problem, these data no longer exhibit a correlation with the mixing
chamber (note that variations are still seen in the mixing chamber; it
is likely that these are not observed at the focal plane itself, where
the NTD thermometry used in the servo loop is performed). There is
minimal correlated low-frequency noise in these 450\,\micron\ data
(note that there is almost no difference once the common-mode has been
subtarcted). The 850\,\micron\ channel, however, exhibits a more
significant signal that is obviously correlated with the telescope
motion. It is also correlated amongst many of the detectors, and
common-mode removal corrects it to some extent. Part of the reason
that this is seen here, and not in the S2SRO data set shown, is that
the amplitude of the scan pattern is larger. Most of this
``scan-synchronous'' noise is attributed to magnetic field pickup, as
described in Section~\ref{sec:magpickup}.

@@

An additional noise source that is significant in only a small subset
of the data (and moreso during the S2SRO period) is magnetic field
pickup. Since the bolometer signals are ultimately detected through
the amplification of magnetic fields, any additional changing fields
within the instrument will also be detected.

@@

Despite the ability of the map-maker to iteratively remove many noise
components, under some circumstances it may be desirable to filter the
data \emph{once} during the pre-processing step. Three main filtering
options are available:

@@

\item the above process can mis-interpret the edges of a bright source
as a step, so we ignore blocks that occur close to bright sources;

@@

\subsubsection{Gap filling / apodization}
\label{sec:gaps}

>>>>

\begin{abstract}
  The Submillimetre Common User Bolometer Array 2 (\scuba) is an
  instrument operating on the 15-m James Clerk Maxwell Telescope,
  consisting of about 5000 bolometers in each of two simultaneous
  imaging bands centered over 450 and 850\,\micron. The camera is
  scanned across the sky and records data at a rate of 200\,Hz. The
  data rate is much greater than that of previously existing
  submillimetre cameras, and represents a significant data analysis
  challenge. We describe the production of \scuba\ maps using the
  Sub-Millimetre User Reduction Facility (SMURF) in which we have
  adopted a fast, iterative approach to map-making that enables data
  reduction on single, modern, high-end desktop computers with
  execution times that are shorter than the observing times.  SMURF is
  used in an automated setting both at the telescope for real-time
  feedback to observers, as well as for the production of generic
  science products at the Canadian Astronomy Data Centre. A suite of
  tools is also provided for interactive data analysis.  Three
  detailed case studies are used to: (i) explore convergence
  properties of the map-maker using simple prior constraints (Uranus
  -- a point source); (ii) achieve the white-noise limit of the
  instrument for faint point-source studies (blank-field survey of the
  Lockman Hole); and (iii) demonstrate that our strategy is capable of
  recovering angular scales up to the size of the array footprint
  (approximately 5\,arcmin) for bright extended sources (star-forming
  region M17).
\end{abstract}

@@

The Submillimetre Common User Bolometer Array 2
\citep[\scuba,][]{holland2012} is a new instrument for the 15-m James
Clerk Maxwell Telescope (JCMT) on Mauna Kea, Hawai'i. The camera
simultaneously images the sky in two broad bands centered over 450 and
850\,\micron, with approximately 7.5 and 14.5\,arcsec full-width at
half-maximum (FWHM) point spread functions (PSFs). The focal planes at
each wavelength are populated with 4 rectangular subarrays, consisting
of $40 \times 32$ bolometers each, and together subtend a nearly
7\,arcmin $\times$ 7\,arcmin footprint on the sky (excluding gaps, the
continuous solid angle is about 43\,arcmin$^{2}$ per focal plane).
This paper describes the properties of \scuba\ data that are relevant
to producing maps of the imaging data, and the Submillimetre User
Reduction Facility, SMURF, a software package for performing the
reduction written using the Starlink Software Environment
\citep{2009ASPC..411..418J}. The details of the instrument design,
performance, and calibration are given in two companion papers:
\citet{holland2012} and \citet{dempsey2012}.

@@

There are numerous ways to attack this map-making problem, both in
terms of the data-collection method, and processing. The two most
important principles to follow in terms of scan strategy are: (i) to
modulate the astronomical signals of interest in such a way that they
appear in the lowest-noise regions of the bolometer noise power
spectrum, i.e., above the $1/f$ knee; and (ii) to provide good
``cross-linking'', in which each portion of the map is sampled on a
range of temporal scales, again, to help distinguish time-varying
noise features from fixed astronomical sources. In the case of \scuba,
(i) is achieved through fast-scanning of the entire telescope (up to
600\,arcsec\,sec$^{-1}$ slew speeds), such that significant drift in
the bolometers due to low-frequency noise occurs more slowly than the
crossing times for the astronomical scales of interest; and (ii) by
offering scan patterns that cross the sky at a wide range of position
angles, ultimately observing each portion of the map at a number of
different times. Such methods are now used by virtually all existing
ground-based submillimetre cameras
\citep[e.g.,][]{glenn1998,weferling2002,wilson2008,kovacs2008b}., in
favour of ``chopping'' methods (where the secondary is quickly moved
to modulate the signal) that were more appropriate for older
instruments that had poorer low-frequency noise performance, and were
only sensitive to modest angular scales

There are three general styles of map-making that are relevant to
reducing bolometer data in the literature. The simplest ``direct
methods'' involve some basic processing of the data to remove as much
noise contamination as possible, (e.g., using baseline removal and
other simple filters), and then re-gridding these cleaned data into a
map. Such was the basic recipe for the reduction of chopped data from
SCUBA-2's predecessor SCUBA
\citep{1998ASPC..145..216J,2000ASPC..216..559J}, and MAMBO
\citep[e.g.,][]{omont2001}, another camera from the same
generation. Generally speaking, such methods are fast, although
depending on the science goals and noise properties of the data,
they may not achieve the best noise performance on the angular scales
of interest. In more recent years a method that has been popular for
reducing data in fields of faint point sources with BOLOCAM
\citep[e.g.,][]{laurent2005}, and its younger sibling AzTEC
\citep[e.g.,][]{scott2008}, is ``principal component analysis'' (PCA)
cleaning. A statistical black-box removes the most correlated
components of the bolometer signals, enabling the detection of
point-sources very close to the theoretical white-noise limits of the
detectors with reasonable computation times when small numbers of
bolometers are involved (hundreds rather than thousands of
detectors). However, PCA cleaning is not a good generic solution for
producing maps of extended structures, since such sources produce
correlated signals amongst many detectors, and are removed by this
procedure.

The best existing map-making strategies for recovering information on
all angular scales are maximum likelihood techniques, in which the
time-series data are expressed as a sampling of the ``true'' map of
the sky plus noise, and then the equation is inverted to estimate the
map as some weighted linear combination of the data that minimizes the
variance.  The first good description of this method appears in
\citet{janssen1992} for the production of maps from the COsmic
Background Explorer (COBE -- the description is relevant despite the
fact that it used a differential radiometer instead of
bolometers). Other descriptions in the CMB literature include
\citet{tegmark1997} and \citet{stompor2002}, and an application to
data from SCUBA is described in \citet{borys2004}.  The downside to
such methods is that they can be both computationally expensive and
have large memory requirements. While for some experimental designs
fast iterative methods for the inversion do exist with minimal memory
usage \citep[e.g.,][]{wright1996}, for the more general map-making
problem involving many detectors, things are significantly complicated
both by the need to measure the cross power-spectra for all unique
pairs of bolometers, and then to perform the inversion itself. The
most promising maximum-likelihood method that may one day be applied
to \scuba\ data is ``SANEPIC'', which successfully reduced maps from
the Balloon-borne Large-Aperture Submillimeter Telescope, involving
hundreds of bolometers, and correctly incorporating inter-bolometer
noise correlations \citep[][]{patanchon2008}.

@@

The third approach adopted here for \scuba\ is a compromise between
the previous two methods. Under the assumption that a significant
portion of the (predominantly low-frequency) non-white noise sources
can be modelled, an iterative solution is obtained for both the
astronomical image and the parameters of the noise model. Since the
remaining (non-modelled) noise sources are assumed to be white, a
single scalar noise may be calculated for all of the data points from
a given bolometer (since the noise at any instant in time is
uncorrelated with others, nor with the data from other bolometers),
greatly simplifying the measurement of noise properties and the
inversion step that is so complicated in the maximum-likelihood
methods. In principle, it should also be possible to recover a wider
dynamic range in angular scales than the simplest ``direct'' methods,
with an increase in calculation time that is merely linear in the
number of iterations needed, while using a comparable amount of
memory. A fast iterative method for handling chopped data from CMB
experiments was described in \citet{wright1996}, and later applied to
chopped SCUBA data by \citet{johnstone2000}, although in each case the
chopping renders the data essentially white without the need to model
additional noise components. More similar to the algorithm described
here was the iterative pipeline recipe for fitting and removing
baseline drifts in SCUBA scan-map data as the astronomical image
estimate improved \citep{1999ASPC..172..171J}. The two closest modern
relatives of SMURF are the Comprehensive Reduction Utility for SHARC-2
\citep[CRUSH,][]{kovacs2008}, and the pipeline developed for the
Bolocam Galactic Plane Survey \citep{aguirre2010}.

A reasonable model for correlated noise in \scuba\ is a single
``common-mode'' signal seen by all of the bolometers. Iterative
estimation and removal of this signal significantly lowers the $1/f$
knee, without compromising structures on angular scales smaller than
the array footprint. Residual independent drifts at lower frequencies
are removed with an iterative high-pass filter. This strategy enables
SMURF to reduce data faster than they are taken, on single high-end
desktop computers, and it has successfully been used as part of a
real-time pipeline offering feedback to observers at the
telescope. The pipeline is also used to generate near-science grade
products using the JCMT Science Archive \citep{2011ASPC..442..203E}
hosted by the Canadian Astronomy Data Centre (CADC). Details of the
pipeline design are given in \citet{gibb2005} and
\citet{2008AN....329..295C}.

This paper is organized in the following way. We first describe the
properties of \scuba\ data in Section~\ref{sec:data} to motivate the
data model that we have adopted. Next, the details of the SMURF
algorithm are given in Section~\ref{sec:algorithm}. The paper is
concluded in Section~\ref{sec:examples} with three detailed test cases
that span the majority of observation types likely to be undertaken
with \scuba, with an emphasis on the mitigation of divergence
problems, and characterizing the output maps: (i) Uranus, a bright,
compact source (Section~\ref{sec:point}); (ii) the Lockman Hole, a
blind survey of faint point-like source (Section~\ref{sec:cosmo}); and
(iii) the star-forming region M17, including bright, extended emission
(Section~\ref{sec:extended}). All of the data analysed in this paper
are publicly available through the CADC \scuba\ raw-data queries page
for the dates and observation numbers given in the
text\footnote{\texttt{http://www3.cadc-ccda.hia-iha.nrc-cnrc.gc.ca/
jcmt/search/scuba2}}. All of the analysis was performed using
\texttt{Starlink} version ... .

@@

In this section we summarize how \scuba\ works,
(Section~\ref{sec:bolos}), give examples of typical bolometers signals
(Section~\ref{sec:bolosignal}), and the impact of magnetic field
pickup (Section~\ref{sec:magpickup}), and finally illustrate the use
of principal component analysis to explore the correlated noise
properties of \scuba\ data (Section~\ref{sec:pca}).

@@

Incoming light passes through a beam-splitter, and then bandpass
filters, providing simultaneous illumination and wavelength definition
of both the 450 and 850\,\micron\ focal planes. Each focal plan is
populated by four ``subarrays'' (labeled s4a--s4d at 450\,\micron, and
s8a--s8d at 850\,\micron), each consisting of 32 columns and 40 rows
of bolometers. The bolometers are thermal absorbers coupled to
superconducting transition-edge sensors (TESs) for thermometry;
temperature variations in the TESs produce changing currents, and
therefore varying magnetic fields.  The changing magnetic fields are
detected and amplified using chains of superconducting quantum
interference devices (SQUIDs), and the larger output currents are
digitized. Each detector has its own SQUID for the first-stage of the
amplification, but the remaining stages occur within a common chain of
SQUIDs for each of the 32 columns. All 40 rows are read-out in
sequence at a row-visit frequency of about 12\,kHz. Such a high sample
rate is unnecessary to produce maps, so the data are re-sampled to
approximately 200\,Hz before writing to disk.  It should also be noted
that there is an additional 41st row of SQUID readouts that are not
connected to TESs. These ``dark SQUIDs'' track non-thermal noise
sources that are common to each column's amplifier chain. The
relationship between the output digitized current, $I$, and the input
power, $P$, or $dI/dP$ is established using flatfield observations
immediately prior to science observations, in which the output signal
is measured throughout a ramp of the pixel heaters (which provide a
known input power); see Section~2.1 in \citet{dempsey2012} and
Section~5.3 in \citet{holland2012}. Finally, the conversion to
astronomical flux units from power involves a correction for
atmospheric extinction \citep[primarily using the JCMT Water Vapour
Monitor to track line-of-sight opacity variations, see Section~3
in][]{dempsey2012}, and the application of a flux conversion factor
(FCF) which is established from regular measurements of calibrators
such as Uranus \citep[Section~5 in][]{dempsey2012}.

@@

We note that the noise performance of \scuba\ has evolved over time. An
initial ``\scuba\ shared-risk observing'' period (S2SRO) took place
during February and March 2010, during which each of the 450 and
850\,\micron\ focal planes were populated with single subarrays, s4a
and s8d respectively. In addition to having significantly fewer
available bolometers than the current fully-commissioned instrument
(and therefore a mapping speed reduced by a factor of $\sim$4),
instabilities in the fridge lead to a large periodic signal with a
period of about 25\,s that was correlated amongst the detectors. After
upgrading and commissioning the instrument, a servo using newly added
thermometers in the focal planes has effectively mitigated this
problem \citep[Section~2.5 in][]{holland2012}. In addition, there were
improvements to the magnetic shielding (reducing magnetic field pickup
as described in Section~\ref{sec:magpickup}), as well as more
effective removal of aliased noise sources (which has reduced both the
presence of line features and the mean white noise level).

In order to reduce the impact of noise on the final maps, it is
essential that an appropriate scan strategy be used. First, it is
clearly advantageous that the astronomical signals appear in the
lowest-noise region of the power spectrum, namely frequencies above
$f_\mathrm{knee}$. This is achieved by scanning the telescope at large
slew rates, which are practically limited to about
600\,arcsec\,s$^{-2}$ for the JCMT in order to have reasonable
turn-around overheads (the amount of time spent changing direction and
speed, over the largest elevation range possible). At this maximum
speed there is a sample every 3.0\,arcsec, or approximately one third
of the 450\,\micron\ diffraction-limited full-width at half-maximum
(FWHM) -- a typical rule-of-thumb for adequately sampling a Gaussian
point spread function (PSF). The scan patterns are designed,
primarily, to mitigate the effects of low-frequency $1/f$ drift. The
simplest scan one might use is a raster ``boustrophedon'' pattern,
i.e., a series of parallel passes across the field. However, such a
scan yields extremely poor performance for structures that are
transverse to the scan direction, since a long time will have elapsed
between each parallel step. In addition, other periodic noise features
may conspire with such a regularly-spaced pattern to produce periodic
structures in the map, or ``scan-synchronous'' noise. Scan strategies
that visit each region of the map on as many different time scales,
and scan angles as possible, offer the best protection against
large-scale noise; such strategies are said to offer good
cross-linking. As a trade-off between truly random scan patterns and
feasibility of implementation taking into consideration both
mechanical and scheduling constraints, \scuba\ presently has two
primary scan modes: PONG, which covers rectangular areas larger than
the array footprint in a closed pattern that ``bounces'' off the edges
at 45\,degree angles (and also usually combined with a rotation
through a number of fixed position angles to create a ``rotating
PONG'' with even better cross-linking); and a ``constant velocity
daisy'' (CV Daisy) for small areas (of order the array footprint), in
which the telescope moves in a circle, whose centre also slowly traces
out a circle. For a more complete description of the \scuba\ observing
modes, see Section~5 in \citet{holland2012}.

\begin{figure*}
\includegraphics[width=0.49\linewidth]{bolos_point_mix_s2sro.pdf}
\includegraphics[width=0.49\linewidth]{bolos_point_mix.pdf}
\caption{A comparison between single bolometer time-series in each of
  the 450 and 850\,\micron\ bands with the mixing chamber temperature
  and azimuth/elevation pointing offsets, before and after upgrading
  the instrument. The grey signals over-plotted in the top two panels
  show the residual time-series after removing the common-mode
  signals.  Left: data taken during the S2SRO period, observation 29
  on 20100313. There is a strong correlation between the bolometers
  and the roughly $\sim$25\,s oscillation in the fridge, but only a
  minor correlation with the telescope motion. The nearly flat
  common-mode subtracted signals show: (i) that most of the
  low-frequency signal is common to all of the bolometers; and (ii)
  the non-correlated, and predominantly white noise at 450\,\micron\
  is significantly larger than at 850\,\micron. Right: data taken with
  the fully-commissioned instrument, observation 38 on
  20111112. Unlike the S2SRO data, there is no strong signal produced
  by variations in the fridge temperature. The 450\,\micron\ data have
  minimal correlated low-frequency noise (as evidenced by the lack of
  a difference between the raw and common-mode subtracted signals),
  although the 850\,\micron\ data show a common signal correlated with
  the telescope motion (note that this scan has a significantly larger
  amplitude than the S2SRO data set), mostly likely caused by magnetic
  field pickup (see Section~\ref{sec:magpickup} and
  Fig.~\ref{fig:magpickup}).}
\label{fig:bolos_mix}
\end{figure*}

@@

Data from the fully commissioned instrument, observation 38 on
20111112, are shown in the right panel of
Fig.~\ref{fig:bolos_mix}. Having solved the fridge oscillation
problem, these data no longer exhibit a correlation with the mixing
chamber (note that variations are still seen in the mixing chamber; it
is likely that these are not observed at the focal plane itself, where
the thermometry used in the servo loop is performed). There is minimal
correlated low-frequency noise in these 450\,\micron\ data (note that
there is almost no difference once the common-mode has been
subtracted). The 850\,\micron\ channel, however, exhibits a more
significant signal that is obviously correlated with the telescope
motion. It is also correlated amongst many of the detectors, and
common-mode removal corrects it to some extent. Part of the reason
that this is seen here, and not in the S2SRO data set shown, is that
the amplitude of the scan pattern is larger. Most of this
``scan-synchronous'' noise is attributed to magnetic field pickup, as
described in Section~\ref{sec:magpickup}.

@@

An additional noise source that is significant in only a small subset
of the data (and more so during the S2SRO period) is magnetic field
pickup. Since the bolometer signals are ultimately detected through
the amplification of magnetic fields, any additional changing fields
within the instrument will also be detected.

@@

Despite the ability of the map-maker to iteratively remove many noise
components, under some circumstances it may be desirable to filter the
data once during the pre-processing step. Three main filtering options
are available:

@@

\item the above process can misinterpret the edges of a bright source
as a step, so we ignore blocks that occur close to bright sources;

@@

\subsubsection{Gap filling / apodisation}
\label{sec:gaps}
